---
title: "Introduction to msigdbeh"
package: msigdbeh
date: "`r doc_date()`"
output: 
  BiocStyle::html_document:
    self_contained: yes
    toc: true
    toc_float: true
    code_folding: show
vignette: >
  %\VignetteIndexEntry{Introduction to msigdbeh}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}  
---

# Overview

The `r Biocpkg("msigdbeh")` R package provides Molecular Signatures Database (MSigDB) gene sets in an R-friendly "tidy" format.
Each row contains one gene (as gene symbols as well as NCBI Entrez and Ensembl IDs) and the corresponding gene set.

The package is designed to serve as the data source for `r CRANpkg("msigdbr")`.

# Installation

The package can be installed from Bioconductor.

```{r install-msigdbeh, eval=FALSE}
if (!require("BiocManager", quietly = TRUE)) {
    install.packages("BiocManager")
}

BiocManager::install("msigdbeh")
```

# Usage

Load `r Biocpkg("msigdbeh")`.

```{r load-msigdbeh, message=FALSE}
library(msigdbeh)
```

Retrieve the latest version of the MSigDB gene sets.
MSigDB is split into human and mouse resources, each one provided in the approved gene symbols of its respective species.

```{r msigdbeh, message=FALSE}
mdf <- msigdbeh(species = "HS")
```

```{r}
mdf[1:10, 1:8]
```

# Dataset summary

Check the included collections.

```{r collections}
mdf[, c("gs_collection", "gs_subcollection", "gs_collection_name")] |>
    sort_by(~ list(gs_collection, gs_subcollection, gs_collection_name)) |>
    unique()
```

Check the database version.

```{r db-version}
unique(mdf$db_version)
```

# ExperimentHub access

The `r Biocpkg("msigdbeh")` package exists primarily for providing documentation.
The relevant data files are stored at the ExperimentHub web service and can be retrieved directly from there.

Load `r Biocpkg("ExperimentHub")` to access resources stored at the ExperimentHub web service.

```{r load-eh, message=FALSE}
library(ExperimentHub)
```

Create an ExperimentHub object.

```{r eh}
eh <- ExperimentHub()
```

Get the data files associated with the `r Biocpkg("msigdbeh")` package using the query method.

```{r query, message=FALSE}
msigdbehData <- query(eh, "msigdbeh")
```

```{r}
msigdbehData
```

A specific object can be downloaded by specifying its identifier.

```{r, message=FALSE}
mdf <- msigdbehData[["EH9632"]]
```

Each file is retrieved from the ExperimentHub server and cached locally.
It should download much more quickly after the initial retrieval.

# Session info

Report information about R, the OS, and attached/loaded packages.

```{r session-info}
sessionInfo()
```
